This is the title [Article v\versionnumber]

### 2022-09-13 manuscript/manuscript.tex
Add further text
Add cheatsheet
Add structure preperation
Remove cleared comments

Recent work has improved the calculation speed tremendously, allowing for the treatment of large biological systems such as proteases or antibodies.
Furthermore, an implementation based on the particle mesh Ewald (PME) \cite{Darden1993-pme} method was presented, which improves the consistency between GIST and popular molecular dynamics engines. 
Yet further changes allow the inclusion of other rigid solvents than water or allow for the treatment of solvent mixtures and ions in solution.\todo{Sources!}

@@

\section{Prerequisites}
\todo{ (vah)  Improve flow and logical progression of prerquisites.}

@@

In the original version of GIST, the energies are calculated based on the the minimum image convention.
In PME-GIST, the electrostatics are calculated using the particle mesh Ewald (PME) method, which yields energies that are highly consistent with the treatment in the Amber MD engine \cite{Chen2021}.
The electrostatic and Lennard-Jones parts $E_\text{elec}$ and $E_\text{lj}$ are computed separately.
\todo{E$_{LJ}$ seems more intuitive to me, need to check what the GIST literature is using -vah. They use $E_{lj}$, but LJ in the text -fwa.}
\begin{equation}
	E_\text{total} = E_\text{elec} + E_\text{lj}
\end{equation}

@@

In \software{ccptraj}, the rectangular prism is defined by parameters of the \inlinecode{gist} command.
The center of the grid is defined by \inlinecode{gridcntr}, the dimensions of the grid is defined by \inlinecode{griddim}, and the spacing is defined by \inlinecode{gridspacn}.

@@

The region can be defined using a ligand in a specific binding pocket of the protein, a specific residue of the protein, or the whole protein itself. A helper script FindCentroid.py \todo{(upload FindCentroid.py!)} can be used to determine the centroid of the region of interest as well as the dimensions of the rectangular prism that will fully enclose the region. A pdb file is needed as an input and the rectangular prism will be centered based on the centroid of the given structure.

@@

\pdfcomment{Franz: I removed the list because it is very long and not complete. We can add it to the lists at the end, but then it should be complete.}
%\todo{Aside from the commands defining the region of interest, there are other options available to specify the GIST analysis. Here is the full list that can be added to the third line of the input file:} 
%\begin{itemize}
%	\item{[doorder] Calculate the water order parameter [reference] for each voxel.}
%	\item{[doeij] Calculate the triangular matrix representing the water-water interactions between pairs of voxels (see below).}
%	\item{[skipE] Skip all energy calculations (cannot be specified with ’doeij’).}
%	\item{[skipS Skip all the entropy calculations.]}
%	\item{[refdens rdval>] Reference density of bulk water, used in computing $g_O, g_H$, and the translational entropy. Default is 0.0334 molecules/Å3.}
%	\item{[temp <tval>] Temperature of the input trajectory.}
%	\item{[noimage] Disable distance imaging in energy calculation.}
%	\item{[gridcntr <xval> <yval> <zval>] Coordinates (Å) of the center of the grid (default 0.0, 0.0, 0.0).}
%	\item{[griddim <xval> <yval> <zval>] Grid dimensions along each coordinate axis (default 40, 40, 40).}
%	\item{[gridspacn <spaceval>] Grid spacing (linear dimension of each voxel) in Angstroms. Values greater than 0.75 Å are not recommended (default 0.5 Å).}
%	\item{[prefix <filename prefix>] Output file name prefix (default “gist”).}
%	\item{[ext <grid extension>] Output grid file name extension (default “.dx”).}
%	\item{[out <output>] Name of the main GIST output file. If not specified set to ’<prefix>-output.dat’.}
%	\item{[info <info>] Name of main GIST info file. If not specified info is written to standard output.}
%\end{itemize}

@@

@@

Here, we choose not to include any sampling of the conformational states. 
It is therefore important that the conformations of the monomers in the complex and dissociated states match exactly. 
Use \software{cpptraj} to produce biotin and streptavidin structures based on the complex by stripping the respective other molecule like this:
\begin{lstlisting}
parm complex/solvated.parm7
trajin complex/equil/EQUIL-DONE.rst
strip ^2 parmout streptavidin/solvated.parm7
trajout streptavidin/pre-equil.ncrst
\end{lstlisting}
(Adapt for biotin as necessary.)
\todo{vah: Maybe include a section on how to prepare a protein structure before equilibration. Protonation, force field and a large enough iso solvation box seem important to me. I will probably also at least mention what simulation settings we are running later on.}
%We provide a script called \software{cpptraj\_remove\_mol.sh} that removes a single molecule from a topology and structure using \software{cpptraj} without altering the rest of the system. \todo{Explain cpptraj rather than our script. The script is for the automated workflow only.}

Use the \software{equilibration.py} script (or your own) to perform short NVT and NpT equilibration runs on the starting structure.
\todo {NPT}
Note that the barostat might interfere with restrained equilibration of the complex, since different molecules are moved relative to each other when the volume changes.
But since our equilibration script also contains unrestrained pressure equilibration, we should not run into any trouble.
\todo{Should we mention more sophisticated choices of structure here, eg clustering? Just taking the "equilibrated" structure is kinda random. -vah}
Then split the equilibrated structure into biotin and streptavidin parts using \software{cpptraj\_remove\_mol.sh}.
After that, you can use \software{equilibration.py -R} to run short restrained equilibrations (where only the water is allowed to move) on the individual systems.

@@

\subsection{Running the GIST analysis for Streptavidin-Biotin Complex}
For easier post-processing, we will use grids that contain the whole molecule.
This could in theory be avoided for streptavidin and the complex, since only the surrounding of the biotin molecule (i.e., the binding pocket) is required for the analysis.
We also recommend that you center the molecule in space.
For each system (biotin, streptavidin, complex), produce a centered pdb file using \software{cpptraj}:
\begin{lstlisting}
parm solvated.parm7
trajin md-01.nc 1 1 1
autoimage !(:WAT) origin
strip :WAT
trajout solute-centered.pdb
\end{lstlisting}
Then, find the extent (minimum and maximum) of the x-, \mbox{y-,} and z-coordinates for each system.
This can be done e.g., using the \software{mdtraj} package in Python.
\begin{lstlisting}[style=python]
import mdtraj as md
pdb = md.load("solute-centered.pdb")
# mdtraj converts to nm.
# *10 to convert back to Angstrom.
min = pdb.xyz[0].min(0) * 10.
max = pdb.xyz[0].max(0) * 10.
print(*list(zip("xyz", min, max)), sep="\n")
\end{lstlisting}
Then, decide on grid dimensions for every system.
For a system centered at the origin with a spacing of \SI{0.5}{\angstrom} and a buffer of, e.g., \SI{15}{\angstrom} between the solute and the grid wall, you need $4.0\times (d+15)$ bins in the respective direction, where $d$ is the larger of $-$min and $+$max in the respective direction.
\todo{this does not work with the default gisttools values for the automatic detection of the energy reference. Better use a larger value. | changed from 7 to 15 -vah }
We note that this centered grid is usually larger than it needs to be in some directions.
If you want to save some extra memory and calculation time, feel free to align the grid more precisely to the minimum and maximum coordinates, or use a smaller grid that only covers the binding pocket.
You can also rotate the molecule by its principal axes to fit the cuboid grid more exactly.
However, this needs to be done \emph{before} the MD run, since rotating the trajectory damages the periodic box information.

@@

\begin{lstlisting}[style=python]
import gisttools as gt
import matplotlib.pyplot as plt
gist = gt.gist.load_gist_file(
    "gist.dat", struct="solute-centered.pdb")
bins, eww = gist.rdf(
    "Eww_unref_norm", bins=100, rmax=20., normalize="norm")
plt.plot(bins, eww)
eww_auto = gist.detect_reference_value().values[0]
plt.axhline(eww_auto, color='k')
plt.gca().set(
    xlabel='distance [A]', ylabel='E_ww per mol')
plt.show()
\end{lstlisting}
\todo{While not massively important, I don't like that we call the gist object here "gist", as that's the same name as gisttools subpackage gist. "gt.gist.bla" and "gist.bla" therefore look quite similiar, but do very different things. -vah}
The expected output of this code is shown in Figure~\ref{fig_ewwref}.

@@

compl = load_gist_file('complex/gist/gist.dat', \
    struct='complex/gist/solute-centered.pdb')
print(compl.n_frames, compl.rho0)
biotin = load_gist_file('biotin/gist/gist.dat', \
    struct='biotin/gist/solute-centered.pdb')
print(biotin.n_frames, biotin.rho0)
strept = load_gist_file('streptavidin/gist/gist.dat', \
    struct='streptavidin/gist/solute-centered.pdb')
print(strept.n_frames, strept.rho0)
\end{lstlisting}
\todo{Maybe add more complete print calls? E.g. print('Biotin: Frames=\{biotin.n\_frames\}, Rho=\{biotin.rho0\} -vah)}
Assign reference energies and check them for plausibility.
\begin{lstlisting}[style=python]
biotin.eww_ref = biotin.detect_reference_value()
print("Biotin:", biotin.eww_ref)
strept.eww_ref = strept.detect_reference_value()
print("Streptavidin:", strept.eww_ref)
compl.eww_ref = compl.detect_reference_value()
print("Complex:", compl.eww_ref)
\end{lstlisting}

@@

\subsection{The CPPTRAJ action}
\subsection{Output}
GIST calculations put out a variety of data sorted by voxel into a file specified by the "out" keyword. Some of the output data is also automatically written to "open data explorer" (.dx) files for convenient visualisation in software such as VMD or PyMOL. The following columns can be found in the output file:
\begin{tabular}{l l p{.3\textwidth}}
\toprule
gX & X & For every element in the main solvent, the number density of atoms found in the voxel, in units of the bulk density. If the same element occurs multiple times, the bulk density is scaled accordingly. \\
Esw & X &  Mean solute-water interaction energy density. \\
Eww & X & Mean water-water interaction energy density. \\
PME & [pme] & Water PME energy density. \\
U\_PME & [pme] & Water PME energy density. \\
dTStrans & X & First order translational entropy density. \\
dTSorient & X & First order orientational entropy density . \\
neighbor & X & Mean number of waters neighboring the water molecules found in this voxel multiplied by the voxel number density. \\
dipole & X & Magnitude of mean dipole moment (polarization). \\
order & X & (if \inlinecode{doorder} was specified) Average Tetrahedral Order Parameter. \\
dipolex & X & x-component of the mean water dipole moment density \\
dipoley & X & y-component of the mean water dipole moment density \\
dipolez & X & z-component of the mean water dipole moment density \\
Eij & X & (if \inlinecode{doeij} was specified) Water-water interaction matrix. \\
\bottomrule
\end{tabular}

\section{Checklists}
%Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
%Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
%One example is shown here.

>>>>

Recent work has improved the calculation speed tremendously, allowing for the treatment of large biological systems such as proteases or antibodies.
Furthermore, an implementation based on the particle mesh Ewald (PME) \cite{Darden1993-pme} method was presented, which improves the consistency between GIST and popular molecular dynamics engines. 
Yet further changes allow the inclusion of other rigid solvents than water or 
allow for the treatment of solvent mixtures and ions in solution. 
\todo{Sources!}

@@

\section{Prerequisites}

@@

In the original version of GIST, the energies are calculated based on the the minimum image convention.
In PME-GIST, the electrostatics are calculated using the particle mesh Ewald (PME) method, which yields energies that are highly consistent with the treatment in the Amber MD engine \cite{Chen2021}.
The electrostatic and Lennard-Jones parts $E_\text{elec}$ and $E_\text{lj}$ are computed separately.

\begin{equation}
	E_\text{total} = E_\text{elec} + E_\text{lj}
\end{equation}

@@

In \software{ccptraj}, the rectangular prism is defined by parameters of the \inlinecode{gist} command.
The center of the grid is defined by \inlinecode{gridcntr}, the dimensions of 
the grid is defined by \inlinecode{griddim} (in voxel counts), and the spacing 
is defined by \inlinecode{gridspacn} (in $\AA$).

@@

The region can be defined using a ligand in a specific binding pocket of the 
protein, a specific residue of the protein, or the whole protein itself. A 
helper script FindCentroid.py can be used to determine the centroid of the 
region of interest as well as the dimensions of the rectangular prism that will 
fully enclose the region. A pdb file is needed as an input and the rectangular 
prism will be centered based on the centroid of the given structure.

@@

@@

The choice of the initial structure to be analyzed is crucial to the results 
you can expect from the GIST calculation. Crystal structures may not be 
representative of the dominant solution structure for a given analysis target. 
In our example we use the final structure after a short equilibration run. A 
more sophisticated approach would be identifying a dominant structure out of 
the ensemble of sampled structures.  Depending on what properties you are 
interested in, running a MD simulation of your structure beforehand and 
choosing a dominant structure or even multiple structures through clustering 
might therefore be beneficial.
When starting with a crystal structure, special care has to be taken during the 
preperation of the structure. Make sure to protonate your structure and take 
care to choose a proper protonation state for the varying chemical moeties in 
your structure. You will need to solvate your structure in a solvent box, which 
should be chosen large enough so that the solvent around the region of interest 
is not influenced through periodic images.

@@

Here, we choose not to include any sampling of the conformational states. 
It is therefore important that the conformations of the monomers in the complex and dissociated states match exactly. 
Use \software{cpptraj} to produce biotin and streptavidin structures based on the complex by stripping the respective other molecule like this:
\begin{lstlisting}
parm complex/solvated.parm7
trajin complex/equil/EQUIL-DONE.rst
strip ^2 parmout streptavidin/solvated.parm7
trajout streptavidin/pre-equil.ncrst
\end{lstlisting}
(Adapt for biotin as necessary.)
%We provide a script called \software{cpptraj\_remove\_mol.sh} that removes a single molecule from a topology and structure using \software{cpptraj} without altering the rest of the system. \todo{Explain cpptraj rather than our script. The script is for the automated workflow only.}

Use the \software{equilibration.py} script (or your own) to perform short NVT and NpT equilibration runs on the starting structure.
\todo {NPT}
Note that the barostat might interfere with restrained equilibration of the complex, since different molecules are moved relative to each other when the volume changes.
But since our equilibration script also contains unrestrained pressure 
equilibration, we should not run into any trouble.

Then split the equilibrated structure into biotin and streptavidin parts using \software{cpptraj\_remove\_mol.sh}.
After that, you can use \software{equilibration.py -R} to run short restrained 
equilibrations (where only the water is allowed to move) on the individual 
systems.

@@

\subsection{Running the GIST analysis for Streptavidin-Biotin Complex}
For easier post-processing, we will use grids that contain the whole molecule.
This could in theory be avoided for streptavidin and the complex, since only the surrounding of the biotin molecule (i.e., the binding pocket) is required for the analysis.
We also recommend that you center the molecule in space.
For each system (biotin, streptavidin, complex), produce a centered pdb file using \software{cpptraj}:
\begin{lstlisting}
parm solvated.parm7
trajin md-01.nc 1 1 1
autoimage !(:WAT) origin
strip :WAT
trajout solute-centered.pdb
\end{lstlisting}
Then, find the extent (minimum and maximum) of the x-, \mbox{y-,} and z-coordinates for each system.
This can be done e.g., using the \software{mdtraj} package in Python.
\begin{lstlisting}[style=python]
import mdtraj as md
pdb = md.load("solute-centered.pdb")
# mdtraj converts to nm.
# *10 to convert back to Angstrom.
min = pdb.xyz[0].min(0) * 10.
max = pdb.xyz[0].max(0) * 10.
print(*list(zip("xyz", min, max)), sep="\n")
\end{lstlisting}
Then, decide on grid dimensions for every system.
For a system centered at the origin with a spacing of \SI{0.5}{\angstrom} and a buffer of, e.g., \SI{15}{\angstrom} between the solute and the grid wall, you need $4.0\times (d+15)$ bins in the respective direction, where $d$ is the larger of $-$min and $+$max in the respective direction.
\todo{ This is handled differently by FindCentroid.py: d = max - min; voxelnr = 
d*5 for x,y,z each; we should handle it similiar or at least mention 
FindCentroid.py here. 
What does gisttools do if the grid is too small? -vah }
We note that this centered grid is usually larger than it needs to be in some directions.
If you want to save some extra memory and calculation time, feel free to align the grid more precisely to the minimum and maximum coordinates, or use a smaller grid that only covers the binding pocket.
You can also rotate the molecule by its principal axes to fit the cuboid grid more exactly.
However, this needs to be done \emph{before} the MD run, since rotating the trajectory damages the periodic box information.

@@

\begin{lstlisting}[style=python]
import gisttools as gt
import matplotlib.pyplot as plt
gist_obj = gt.gist.load_gist_file(
    "gist.dat", struct="solute-centered.pdb")
bins, eww = gist_obj.rdf(
    "Eww_unref_norm", bins=100, rmax=20., normalize="norm")
plt.plot(bins, eww)
eww_auto = gist_obj.detect_reference_value().values[0]
plt.axhline(eww_auto, color='k')
plt.gca().set(
    xlabel='distance [A]', ylabel='E_ww per mol')
plt.show()
\end{lstlisting}

The expected output of this code is shown in Figure~\ref{fig_ewwref}.

@@

compl = load_gist_file('complex/gist/gist.dat', \
    struct='complex/gist/solute-centered.pdb')
print(compl.n_frames, compl.rho0)
biotin = load_gist_file('biotin/gist/gist.dat', \
    struct='biotin/gist/solute-centered.pdb')
print(biotin.n_frames, biotin.rho0)
strept = load_gist_file('streptavidin/gist/gist.dat', \
    struct='streptavidin/gist/solute-centered.pdb')
print(strept.n_frames, strept.rho0)
\end{lstlisting}
Assign reference energies and check them for plausibility.
\begin{lstlisting}[style=python]
biotin.eww_ref = biotin.detect_reference_value()
print("Biotin:", biotin.eww_ref)
strept.eww_ref = strept.detect_reference_value()
print("Streptavidin:", strept.eww_ref)
compl.eww_ref = compl.detect_reference_value()
print("Complex:", compl.eww_ref)
\end{lstlisting}

@@

\begin{Checklists*}

\begin{checklist}{The CPPTRAJ action}

\textbf{Options}\\
Various flags and options can be provided when running a GIST calculation in 
cpptraj. A list of possible and required options is provided here:

\begin{tabular}{l p{.7\textwidth}}
\toprule
doorder & Calculate the water order parameter [reference] for each voxel\\
doeij & Calculate the triangular matrix representing the water-water 
interactions between pairs of voxels.\\
skipE & Skip all energy calculations (cannot be specified with ’doeij’)\\
skipS & Skip all the entropy calculations.\\
refdens <rdval> & Reference density of bulk water, used in computing 
$g_O, g_H$, and the translational entropy. Default is 0.0334 molecules/Å3. \\
temp <tval> & Temperature of the input trajectory. \\
noimage &  Disable distance imaging in energy calculation. \\
gridcntr <xval> <yval> <zval> & Coordinates (Å) of the center of the grid 
(default 0.0, 0.0, 0.0). \\
griddim <xval> <yval> <zval> & Grid dimensions along each coordinate axis 
(default 40, 40, 40).\\
gridspacn <spaceval> & Grid spacing (linear dimension of each voxel) in 
Angstroms. Values greater than 0.75 Å are not recommended (default 0.5 Å).\\
prefix <filename prefix> & Output file name prefix (default “gist”).\\
ext <grid extension> & Output grid file name extension (default “.dx”). \\
out <output> & Name of the main GIST output file. If not specified set to 
’<prefix>-output.dat’.\\
info <info> & Name of main GIST info file. If not specified info is 
written to standard output. \\
\bottomrule
\end{tabular}

\textbf{Output}\\
GIST calculations put out a variety of data sorted by voxel into a file 
specified by the "out" keyword. Some of the output data is also automatically 
written to "open data explorer" (.dx) files for convenient visualisation in 
software such as VMD or PyMOL. The following columns can be found in the output 
file:
\begin{tabular}{l l p{.8\textwidth}}
\toprule
Name & Keyword & Description\\
\midrule
gX & X & For every element in the main solvent, the number density of atoms found in the voxel, in units of the bulk density. If the same element occurs multiple times, the bulk density is scaled accordingly. \\
Esw & X &  Mean solute-water interaction energy density. \\
Eww & X & Mean water-water interaction energy density. \\
PME & [pme] & Water PME energy density. \\
U\_PME & [pme] & Water PME energy density. \\
dTStrans & X & First order translational entropy density. \\
dTSorient & X & First order orientational entropy density . \\
neighbor & X & Mean number of waters neighboring the water molecules found in this voxel multiplied by the voxel number density. \\
dipole & X & Magnitude of mean dipole moment (polarization). \\
order & X & (if \inlinecode{doorder} was specified) Average Tetrahedral Order Parameter. \\
dipolex & X & x-component of the mean water dipole moment density \\
dipoley & X & y-component of the mean water dipole moment density \\
dipolez & X & z-component of the mean water dipole moment density \\
Eij & X & (if \inlinecode{doeij} was specified) Water-water interaction matrix. \\
\bottomrule
\end{tabular}
\end{checklist}
\end{Checklists*}
\section{Checklists}
%Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
%Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
%One example is shown here.
